Hydra Battles and Cie

### 2021-08-06 doc/part-hydras.tex
converting chapter 3 to alectryon (continued)
converting chapter 3 to alectryon (draft, completed)

\label{sect:on-lt-notation}
\label{sect:on-le-notation}
\label{sect:measure-ON}

@@

\label{def: Finite-ord-type}
\begin{Coqsrc}
Coercion is_true: bool >-> Sortclass.

Definition t (n:nat) := {i:nat | Nat.ltb i n}.
\end{Coqsrc}

The order on type (\texttt{t $n$}) is defined through the projection on \texttt{nat}.

\begin{Coqsrc}
Definition lt {n:nat} : relation (t n) :=
  fun alpha beta => Nat.ltb ( proj1_sig alpha) (proj1_sig beta).
\end{Coqsrc}

@@

\begin{Coqsrc}
Program Example alpha1 : t 7 := 2.

Program Example beta1 : t 7 := 5.

Example i1 : lt  alpha1 beta1.
Proof. now compute. Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma t0_empty (alpha: t 0): False.
Proof.
  destruct alpha; discriminate.
Qed.

Program Definition bad : t 10 := 10.
Next Obligation.
  compute.
\end{Coqsrc}

\begin{Coqanswer}
1 subgoal (ID 162)

  ============================
  false = true
\end{Coqanswer}

\begin{Coqsrc}
Abort.
\end{Coqsrc}

Note also that attempting to compare a term  of type (\texttt{t $n$}) with a term of
type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.

\begin{Coqsrc}
Program Example gamma1 : t 8 := 7.

Fail Goal lt alpha1 gamma1.
\end{Coqsrc}

\begin{Coqanswer}
 The command has indeed failed with message:
The term "gamma1" has type "t 8" while it is expected to have type "t 7".
\end{Coqanswer}

@@

\begin{Coqsrc}
Definition compare {n:nat} (alpha beta : t n) :=
  Nat.compare (proj1_sig alpha) (proj1_sig beta).

Lemma compare_correct {n} (alpha beta : t n) :
  CompareSpec (alpha = beta) (lt alpha beta) (lt beta alpha)
              (compare alpha beta).
\end{Coqsrc}

@@

@@

\begin{Coqsrc}
Lemma lt_wf (n:nat) : well_founded (@lt n).
\end{Coqsrc}

@@

\begin{Coqsrc}
#[global] Instance sto n : StrictOrder (@lt n).

#[global] Instance comp n: Comparable (@lt n) compare.

#[global] Instance FinOrd n : ON (@lt n) compare. 
Proof.
  split.
  - exact (comp n).
  - apply lt_wf.
Qed.
\end{Coqsrc}

\begin{remark}
It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}. In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$. 
In our formalization, the interpretation of an ordinal as a set is realized by the following definition
(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).

\begin{Coqsrc}
Definition bigO `{nA : @OrdinalNotation A ltA stoA compareA}
           (a: A) : Ensemble A :=
  fun x: A => ltA x a.
\end{Coqsrc}
\end{remark}

@@

\begin{Coqsrc}
Class  SubON 
       `(OA : @ON A ltA  compareA)
       `(OB : @ON B ltB  compareB)
       (alpha :  B)
       (iota : A -> B):=
  {
  SubON_compare: forall x y : A,  compareB (iota x) (iota y) =
                                 compareA x y;
  SubON_incl : forall x, ltB (iota x) alpha;
  SubON_onto : forall y, ltB y alpha  -> exists x:A, iota x = y}.
\end{Coqsrc}

@@

\begin{Coqsrc}
Instance Incl : SubON Omega Omega_plus_Omega omega fin.
\end{Coqsrc}

@@

We are now able to build an instance of \texttt{SubON}.

@@

\begin{Coqsrc}
Section Inclusion_ij.

  Variables i j : nat.
  Hypothesis Hij : (i < j)%nat.

  Remark Ltb_ij : Nat.ltb i j.

  Program Definition iota_ij  (alpha: t i) : t j :=  alpha.

   Let b : t j := exist _ i Ltb_ij.

   Global Instance F_incl_ij  : SubON  (FinOrd i) (FinOrd j) b iota_ij.
  (* ... *)

  End Inclusion_ij.
\end{Coqsrc}

@@

\begin{Coqsrc}
Class ON_correct `(alpha : Ord)
     `(OA : @ON A ltA compareA)
      (iota : A -> Ord) :=
  { ON_correct_inj : forall a, lt (iota a) alpha;
    ON_correct_onto : forall beta, lt beta alpha ->
                                exists b, iota b = beta;
    On_compare_spec : forall a b:A,
        match compareA a b with
          Datatypes.Lt => lt (iota a) (iota b)
        | Datatypes.Eq => iota a = iota b
        | Datatypes.Gt => lt (iota b) (iota a)
        end
  }.
\end{Coqsrc}

@@

@@

\label{types:ON-iso} 
\begin{Coqsrc}
Class  ON_Iso 
       `(OA : @ON A ltA compareA)
       `(OB : @ON B ltB  compareB)
       (f : A -> B)
       (g : B -> A):=
  {
  iso_compare: forall x y : A, 
      compareB (f x) (f y) = compareA x y;
  iso_inv1 : forall a, g (f a)= a;
  iso_inv2 : forall b, f (g b) = b
}.
\end{Coqsrc}

>>>>

\label{sect:on-lt-notation}
\label{sect:on-le-notation}
\label{sect:measure-ON}
\label{sect:bigO-ON}

@@

\label{def: Finite-ord-type}

\input{movies/snippets/ON_Finite/Defs}

@@

\input{movies/snippets/ON_Finite/Example1}

@@

 \input{movies/snippets/ON_Finite/t0Empty}

\begin{Coqsrc}
Program Definition bad : t 10 := 10.
Next Obligation.
\end{Coqsrc}

\begin{Coqanswer}
1 subgoal (ID 118)

  ============================
  10 <? 10 
\end{Coqanswer}

\begin{Coqsrc}
  compute.
Abort.
\end{Coqsrc}

%\input{movies/snippets/ON_Finite/bad}

@@

\input{movies/snippets/ON_Finite/compareDef}

@@

Please note  that attempting to compare a term  of type (\texttt{t $n$}) with a term of
type (\texttt{t $p$})  leads to an error if $n$ and $p$ are not convertible.

\input{movies/snippets/ON_Finite/Example2}

@@

\input{movies/snippets/ON_Finite/ltWf}

@@

\input{movies/snippets/ON_Finite/ONInstance}

\begin{remark}
It is important to keep in mind  that the integer $n$ is not an ``element'' of \texttt{FinOrd $n$}. In set-theoretic presentations of ordinals, the set associated with the ordinal $n$ is $\{0,1,\dots,n-1\}$. 
In our formalization, the interpretation of an ordinal as a set is realized by the function \texttt{bigO} (see Section\vref{sect:bigO-ON}).
\end{remark}

@@

following definition
(in ~\href{../theories/html/hydras.OrdinalNotations.ON_Generic.html}{ON\_Generic}).

\input{movies/snippets/ON_Generic/SubONDef}

@@

\input{movies/snippets/ON_Omega_plus_omega/Incl}

@@

@@

\input{movies/snippets/ON_Finite/InclIJ}
\input{movies/snippets/ON_Finite/InclIJa}
\input{movies/snippets/ON_Finite/InclIJb}
\input{movies/snippets/ON_Finite/InclIJc}
\input{movies/snippets/ON_Finite/InclIJd}

@@

\input{movies/snippets/ON_Generic/ONCorrect}

@@

\vspace{4pt}
\noindent\emph{From Module~\href{../theories/html/hydras.Schutte.Correctness_E0.html}{Schutte.Correctness\_E0}}

\input{movies/snippets/Correctness_E0/Epsilon0Correct}

@@

\label{types:ON-iso}

\input{movies/snippets/ON_Generic/ONIso}
